Pyg4ometry: a python library for the creation of Monte Carlo radiation transport physical geometries

### 2020-01-16 public/paper/pyg4ometry.tex

\subsection{Visualisation}
A  geometry hierarchy can be viewed using the popular Visualisation Toolkit (VTK). No separate scene graph is required as the Geant4 volume hierarchy is sufficient 
to place the meshes associated with each physical volume. A daughter volume is placed within a logical volume with a rotation $\mathbf{R}$, scale $\mathbf{S}$ and 
translation $\mathbf{T}$.

@@

\section{Geant4 to FLUKA conversion}
It is relatively straight forward to convert Geant4 geometry to FLUKA. Each of the Geant4 solids can be mapped to a FLUKA CSG region.
Some solids in Geant4 directly map to a FLUKA body, others need a construction of a simple FLUKA CSG tree.

\begin{table}[hbt!]
\centering
\begin{tabular}{| l | l | } \hline
Geant4 solid			& FLUKA region constuction		\\ \hline
Box					& +2 XYP + 2 XZP + 2 YZP 		\\
Tube					& +ZCC - 2 XYP	 			\\
CutTube				& +ZCC - 2 PLA				\\
Cone				& +TRC -TRC - WED			\\
Para					& +ARB						\\
Trd					& +ARB						\\
Trap					& +ARB						\\
Sphere				& +SPH - SPH  - WED - TRC		\\
Orb					& +SPH						\\
Torus				& +$N$ ZCC  - $N$ PLA			\\
Polycone				& +$N$ ARB					\\
GenericPolycone		& +$N$ ARB					\\
Polyhedra				& +$N$ ARB					\\
Eltube				& +ZEC  - 2 XYP				\\
Ellipsoid				& +ELL - 2 XYP		 			\\
Elcone				& +QUA - 2 XYP				\\
Paraboloid			& +QUA - 2 XYP				\\
Hype					& +QUA - 2 XYP				\\
Tet					& +ARB						\\
Xtru					& +$N$ ARB					\\
TwistedBox			& +$N$ ARB					\\
TwistedTtap			& +$N$ ARB					\\
TwistedTrd			& +$N$ ARB				 	\\
TwistedTube			& +$N$ ARB					\\
Arb8					& +ARB						\\
Tessellated			& N/A					 	\\
Union				& | zone1 | zone2				\\
Subtraction			& +zone1 -zone2				\\
Intersection			& +zone1 +zone2				\\
MultiUnion			& | zone1 | zone2 | zone3 | ...	\\
ScaledSolid			& geant4.solid.ScaledSolid		\\ \hline				
\end{tabular}
\caption{GDML/Geant4 solids and the mapping to FLUKA bodies}
\end{table}

>>>>

\subsection{Visualisation} \label{sec:visualisation}
A  geometry hierarchy can be viewed using the popular Visualisation Toolkit (VTK). No separate scene graph is required as the Geant4 volume hierarchy is sufficient 
to place the meshes associated with each physical volume. A daughter volume is placed within a logical volume with a rotation $\mathbf{R}$, scale $\mathbf{S}$ and 
translation $\mathbf{T}$.

@@

\section{Geant4 to FLUKA conversion}
It is relatively straight forward to convert Geant4 geometry to FLUKA. Each of the Geant4 solids can be mapped to a FLUKA CSG region.
Some solids in Geant4 directly map to a FLUKA body, others need a construction of a simple FLUKA CSG tree. Table~\ref{tab:geant2fluka}
lists the mapping between geant4 solids and the bodies used to compose a fluka region.

\begin{table}[hbt!]
\centering
\begin{tabular}{| l | l | } \hline
Geant4 solid			& FLUKA region constuction		\\ \hline
Box					& +2 XYP + 2 XZP + 2 YZP 		\\
Tube					& +ZCC - 2 PLA -2 XYP - ZCC	 	\\
CutTube				& +ZCC - 4 PLA -ZCC			\\
Cone				& +TRC - TRC - WED			\\
Para					& +ARB						\\
Trd					& +ARB						\\
Trap					& +ARB						\\
Sphere				& +SPH - SPH  - WED - TRC		\\
Orb					& +SPH						\\
Torus				& +$N$ ZCC  - $N$ PLA			\\
Polycone				& +$N$ ARB					\\
GenericPolycone		& +$N$ ARB					\\
Polyhedra				& +$N$ ARB					\\
Eltube				& +ZEC  - 2 XYP				\\
Ellipsoid				& +ELL - 2 XYP		 			\\
Elcone				& +QUA - 2 XYP				\\
Paraboloid			& +QUA - 2 XYP				\\
Hype					& +QUA - 2 XYP				\\
Tet					& +ARB						\\
Xtru					& +$N_{\rm plane} N_{\rm point}$ PLA					\\
TwistedBox			& +$N$ ARB					\\
TwistedTtap			& +$N$ ARB					\\
TwistedTrd			& +$N$ ARB				 	\\
TwistedTube			& +$N$ ARB					\\
Arb8					& +ARB						\\
Tessellated			& N/A					 	\\
Union				& $R_1 \cup R_2$				\\
Subtraction			& $R_1 - R_2$					\\
Intersection			& $R_1 \cap R_2$				\\
MultiUnion			& $R_1 \cup R_2 \cup R_3 \cup$	\\
ScaledSolid			& geant4.solid.ScaledSolid		\\ \hline				
\end{tabular}
\label{tab:geant2fluka}
\caption{GDML/Geant4 solids and the mapping to FLUKA bodies}
\end{table}

There are some important issues which must be considered to perform an accurate conversion. 
A Geant4 logical volume solid might need to be created from a fluka region composing of many zones, 
so $R_1 = +z_1 | +z_2$, similarly a void is required to locate daughter volume placements which again 
might also be converted to a region of multiple zones $R_2= +z_3| z_4$. So creating the fluka region is then 
\begin{eqnarray}
R_1 - R_2 	& = & (+z_1 | +z_2) - ( +z_3 | +z_4) 			\\
			& = & (+z_1 - z_3 - z_4)  | (+z_2 - z_3 - z_4)	
\label{eqn:setDiff}
\end{eqnarray}
Geant4 has the boolean solids associated with difference, union and intersection, so in addition to Equation~\ref{eqn:setDiff}, both $R_1 \cup R_2$ and 
$R_1 \cap R_2$ is required in fluka notation, so 
\begin{eqnarray}
R_1 \cup R_2 	& = & (+z_1 | +z_2)  \cup ( +z_3 | +z_4) \\
			& = & +z_1 | +z_2 |  +z_3 | +z_4
\end{eqnarray}
and 
\begin{eqnarray}
R_1 \cap R_2 	& = & (+z_1 | +z_2) \cap ( +z_3 | +z_4) \\
			& = & +z_1 +z_3  | +z_1 +z_4 | +z_2 +z_3 | +z_2 +z_4
\end{eqnarray}

Fluka (apart from the {\it lattice} directive) has no sense of a volume hierarchy. Each body is placed with a translation, rotation, expansion geometry 
directives in global coordinates. A transformation from world coordinates to a physical volume is built up by applying recursively daughter volume transformations and 
this is used to place Fluka bodies. This  in practice is very similar to the procedure to create the VTK visualisation already described in Section~\ref{sec:visualisation}
